\LaTeX\ Author Guidelines for ICCV Proceedings

### 2021-02-08 chapters/3_algorithm.tex
Polish definition of binary trees and UM

We consider an undirected simple edge-weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. In computer vision applications, the nodes can represent either pixels, superpixels or voxels. We call the set $\Pi$ a \emph{clustering} or \emph{partitioning} with $K$ clusters if $V = \cup_{S\in\Pi} S $, $\,S \cap S' = \emptyset$ for different clusters $S, S'\in \Pi$ and every cluster $S \in \Pi$ induces a connected subgraph of $\mathcal{G}$. We also denote as $S_u$ the cluster associated with node $u$.
The weight function $w^+: E \rightarrow \mathbb{R}^+$ associates to every edge a positive scalar attribute $w_e^+\in \mathbb{R}^+$ representing a merge affinity or a similarity measure: the higher this number, the higher the inclination of the two incident vertices to be assigned to the same cluster\footnote{Note that other formalisms for positively weighted graphs associate distances to the edges, thus, the \emph{lower} the edge weight, the higher the attraction between the two linked nodes, contrary to our definition of $w^+$.}. On the other hand, $w^-: E \rightarrow \mathbb{R}^+$ associates to each edge a split tendency $w_e^- \in \mathbb{R}^+$: the higher this weight, the more the incident vertices would like to be in different clusters. 
Graphs of the type $\mathcal{G}(V,E,w^+, w^-)$ are also often defined as \emph{signed graphs} $\mathcal{G}(V,E,\cost)$, featuring positive and negative edge weights $\cost_e\in \mathbb{R}$. Following the theoretical considerations in \cite{lange2018partial}, we define these signed weights as ${\cost_e = w_e^+ - w_e^-}$. Some approaches directly compute $\cost_e$, whereas others compute $w_e^+$ and $w_e^-$ separately.
In this formalism, graphs with purely attractive interactions are a special case of $\mathcal{G}(V,E,\cost)$ with $\cost_e \geq 0, \, \forall e \in E$. \TODO{remove and add multicut energy, define HC}

@@

\textbf{Hierarchical trees} In HC literature, a \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} whose leaves are in one-to-one correspondence with $V$, representing the merging order of an agglomerative algorithm.
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor of $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
\TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
For each pairs of leaves $u,v$, we can then assign a \emph{height} to their least common ancestor in the tree $T^*$ corresponding to the linkage/interaction between clusters $S_u$ and $S_v$ when $u$ and $v$ were merged by the algorithm:
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T_{\mathrm{L}}[u \vee v]) \big)
\end{equation}

\textbf{Ultrametrics}  Let us define the function $d_T: V \times V \rightarrow \mathbb{R}^{+}$ as:
\begin{align}
d_T(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v').
\end{align}
In the next section, we will show that for some specific choices of linkage $\interact{}$, then $(V, d_{T^*})$ is an ultrametric on the original graph, where $T^*$ is the hierarchical tree built by the \algname{} algorithm.

\begin{itemize}
\item Def: A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.
\item The connection between hierarchical trees and ultrametrics is well known. Normally, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. In our case, the interaction $\treeHeight{}$ can also be negative, so we first map these interactions to strictly positive values:

\item By definition, we have that $d_T(u,v)\geq 0$, $d_T(u,v)=d_T(v,u)$, and $d_T(u,u)= 0$. So, to prove that $(V, d_{T^*})$ is an ultrametric, we need to prove that
\begin{equation}
\interact{}_{T^*}(u,v) \geq \min \{\interact{}_{T^*}(u,t), \interact{}_{T^*}(v,t)\}
\end{equation}
 $\forall u,v,t\in V$. In other words, we want to make sure that, at every iteration, the algorithm merges pairs of clusters with lower and lower interactions. 
\begin{equation}
\interact{}_{T^*}(u,v) \geq \interact{}_{T^*}(u,t) \Leftrightarrow T[u \vee t] \subseteq T[u \vee v] 
\end{equation}
This does not happen for all linkages (see Sum fig. 1) and when constraints are in place.  
 % that have the most positive interaction   two clusters merged by the algorithm will always have an interaction that is more negative as compared to all clusters already merged and more positive than all clusters that will be merged at later iterations of the algorithm.
\item We call a linkage invariant to $\alpha$-shift of weights, if by changing all edge weights of the graphs by a constant $w_e+\alpha$ the structure of the hierarchical tree  does not change and defines the same function $d_T$. (And correlation clustering problem is not like that).
\end{itemize}

@@

In Algorithm \ref{main_alg}, we provide simplified pseudo-code for the proposed \algname{} algorithm. \algname{} implements a bottom-up approach that starts by assigning each node to its own cluster and then iteratively merges pairs of adjacent clusters. The algorithm has two variants. The first one (option \emph{addCannotLinkConstraints} is \texttt{False}) starts by merging clusters with the strongest attractive interaction and stops once the remaining clusters share only mutual repulsive interactions (see iterations on toy graphs in block 4 of Fig.\ \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}}). After each merging iteration, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S_u, S_v)$ listed in Table \ref{tab:linkage-criteria}.

@@

@@

Our main contribution is the generalized algorithm for signed graph partitioning, GASP, that encompasses several known and new agglomerative algorithms on display in Table \ref{tab:linkage-criteria}.
In our framework, individual algorithms are differentiated by the linkage criterion employed. We review them in the following paragraphs.

@@

\begin{figure}[t]
\centering
        \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/example_no_constr.pdf}
        \caption{No constraints}\label{subfig:no_constraints}
    \end{subfigure} \hfill \vspace{8pt}
    \begin{subfigure}[t]{0.46 \textwidth}
        \centering
        \includegraphics[width=\textwidth]{figs/example_with_constr.pdf}
        \caption{With cannot-link constraints}\label{subfig:with_constraints}
    \end{subfigure}
\caption{Some iterations of the generalized algorithm (using \emph{Sum} linkage criteria) with and without adding cannot-link constraints. The graph has both attractive (green) and repulsive (red) edges and cannot-link constraints are shown with triple violet bars on the edges. The edge selected at each iteration is highlighted in yellow. We note that when constraints are enforced, the final clustering is given by two clusters instead of only one.}
\label{fig:algorithm_with_without_CLC}
\end{figure}

@@

            \thead{\textbf{Unsigned Graphs}} & \thead{Sum Linkage\\Hier. Aggl. Clust.} &\thead{Single Linkage\\Hier. Aggl. Clust.} &\thead{Average Linkage\\Hier. Aggl. Clust.} &\thead{Single Linkage\\Hier. Aggl. Clust.} &\thead{Complete Linkage\\Hier. Aggl. Clust.} \\
            \thead{\textbf{Signed Graphs} \\\textbf{Without Constraints}} & \thead{GAEC \cite{keuper2015efficient}} & \thead{Mutex Watershed \cite{wolf2018mutex}} & \thead{Average Linkage\\Hier. Aggl. Clust.} &\thead{Single Linkage\\Hier. Aggl. Clust.} &\thead{Complete Linkage\\Hier. Aggl. Clust.} \\
            \thead{\textbf{Signed Graphs} \\\textbf{With Constraints}} & \thead{\textbf{NEW}} % \thead{Greedy Fixation \cite{levinkov2017comparative}} 
            & \thead{Mutex Watershed \cite{wolf2018mutex}} & \thead{\textbf{NEW}} &  \thead{\textbf{NEW}} &  \thead{\textbf{NEW}} \\
            % \multicolumn{2}{c|}{\multirow{2}{*}[-0.5em]{\thead{\textbf{\algname{} linkage criteria} $\,\,\interact(S_u ,S_v)$}}}  & \multirow{2}{*}[-0.5em]{\thead{\textbf{Unsigned Graphs}}} & \multicolumn{2}{c}{\thead{\textbf{Signed Graphs}}}  \\        
            % \multicolumn{2}{c|}{} &  &  \multicolumn{1}{c}{\thead{No Constraints}} & \thead{With Constraints} \\ \midrule

@@

        \end{tabular}
    \end{subtable} 
    \caption{Existing and new clustering algorithms that can be reformulated as special cases of the proposed generalized algorithm for signed graph partitioning, \algname{}, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. The set $E_{uv}$ is defined as the set of all edges connecting cluster $S_u$ to cluster $S_v$, i.e. $E_{uv}=(S_u \times S_{v \neq u}) \cap E$.}
    \label{tab:linkage-criteria}
\end{table*}

>>>>

We consider an undirected simple edge-weighted graph $\mathcal{G}(V,E,w^+, w^-)$ with both attractive and repulsive edge attributes. In computer vision applications, the nodes can represent either pixels, superpixels or voxels. We call the set $\Pi$ a \emph{clustering} or \emph{partitioning} with $K$ clusters if $V = \cup_{S\in\Pi} S $, $\,S \cap S' = \emptyset$ for different clusters $S, S'\in \Pi$ and every cluster $S \in \Pi$ induces a connected subgraph of $\mathcal{G}$. We also denote as $S_u$ the cluster associated with node $u$.
The weight function $w^+: E \rightarrow \mathbb{R}^+$ associates to every edge a positive scalar attribute $w_e^+\in \mathbb{R}^+$ representing a merge affinity or a similarity measure: the higher this number, the higher the inclination of the two incident vertices to be assigned to the same cluster\footnote{Note that other formalisms for positively weighted graphs associate distances to the edges, thus, the \emph{lower} the edge weight, the higher the attraction between the two linked nodes, contrary to our definition of $w^+$.}. On the other hand, $w^-: E \rightarrow \mathbb{R}^+$ associates to each edge a split tendency $w_e^- \in \mathbb{R}^+$: the higher this weight, the more the incident vertices would like to be in different clusters. 
Graphs of the type $\mathcal{G}(V,E,w^+, w^-)$ are also often defined as \emph{signed graphs} $\mathcal{G}(V,E,\cost)$, featuring positive and negative edge weights $\cost_e\in \mathbb{R}$. Following the theoretical considerations in \cite{lange2018partial}, we define these signed weights as ${\cost_e = w_e^+ - w_e^-}$. Some approaches directly compute $\cost_e$, whereas others compute $w_e^+$ and $w_e^-$ separately.
In this formalism, graphs with purely attractive interactions are a special case of $\mathcal{G}(V,E,\cost)$ with $\cost_e \geq 0, \, \forall e \in E$. \TODO{shorten and add multicut energy def, define HC abbrv.}

\begin{itemize}
\item Given a linkage $\interact{}$, we say that Algorithm \ref{main_alg} is invariant to a $\alpha$ weight-shift, if by changing all edge weights of the graphs by a constant $w_e+\alpha$ the structure of the hierarchical tree  does not change and defines the same function $d_T$.
\item \TODO{Why do we care?} MC does not return nested/hierarchical solutions when shifting weights. Which of these algorithms behave the same, or, somehow are more related to correlation clustering? So, actually, we care more about the shifted-solutions being nested, and not really about the fact that the hierarchy tree built by the algorithm stays the same. Is this somehow equivalent...?
\end{itemize}

@@

@@

\TODO{Update description to new pseudocode} In Algorithm \ref{main_alg}, we provide simplified pseudo-code for the proposed \algname{} algorithm. \algname{} implements a bottom-up approach that starts by assigning each node to its own cluster and then iteratively merges pairs of adjacent clusters. The algorithm has two variants. The first one (option \emph{addCannotLinkConstraints} is \texttt{False}) starts by merging clusters with the strongest attractive interaction and stops once the remaining clusters share only mutual repulsive interactions (see iterations on toy graphs in block 4 of Fig.\ \hyperref[fig:intro_figure]{\ref*{fig:intro_figure}}). After each merging iteration, the interaction between the merged cluster and its neighbors is updated according to one of the linkage criteria $\interact(S_u, S_v)$ listed in Table \ref{tab:linkage-criteria}.

@@

\paragraph{Hierarchical trees and ultrametrics} In HC literature,  a \emph{dendrogram} $T$ is a rooted binary tree\footnote{In general, one could look at trees that are not binary. However, the algorithms discussed in this paper always generate binary hierarchical trees, so nothing would be gained by this generalization.} representing the merging order of an agglomerative algorithm, such that the leaves of this tree are in one-to-one correspondence with $V$. 
Let $T_{\mathrm{R}},T_{\mathrm{L}}\subset T$ denote the subtrees rooted at the two children of the root node in $T$.
For any two leaves $u,v \in V$, let $T[u \vee v]$ be the subtree rooted at the least common ancestor of $u$ and $v$ (furthest from the root), and let \texttt{leaves}$(T[u \vee v])\subseteq V$ be the set of leaves of this subtree. 
% \TODO{Here we talk about the algorithm already, so we could move it to next section and the flow could be better. But we need a tree in the alg. So we better do this later}
Algorithm \ref{main_alg} returns a binary tree denoted by $T^*$.
For each pair of distinct leaves $u,v$, we can then assign a \emph{depth} \TODO{better name?} to their least common ancestor node in $T^*$ defined by the interaction between clusters $S_u$ and $S_v$, when $u$ and $v$ were merged: \TODO{See trees on toy graphs in Fig. 1} 
% Then, given a binary tree $T$ and a linkage criterion $\interact$, we can associate a \emph{signed similarity} function $\treeHeight$ to each pair of distinct original nodes $u\neq v$, with $u, v \in V$:
\begin{equation}
\treeHeight(u,v) \equiv \interact{} \big( \text{\texttt{leaves}}(T^*_{\mathrm{R}}[u \vee v]), \text{\texttt{leaves}}(T^*_{\mathrm{L}}[u \vee v]) \big)
\end{equation}

The connection between hierarchical trees and ultrametrics is well known\footnote{A metric space $(X,d)$ is an \emph{ultrametric} if for every $x,y,z \in X$, $d(x,y)\leq \max \{d(x,z), d(y,z)\}$.}. 
We now want to use the \emph{depth-measure} $\treeHeight{}$ to define an ultrametric on the graph.  
Normally, ultrametrics are associated to strictly positive \emph{similarity} or \emph{dissimilarity} measures on a graph. Thus, we first need to map $\treeHeight{}$ to positive values:
\begin{align}
d_{T^*}(u,v)\equiv &\begin{cases}
0 & \text{if}\,\, u=v\\
M-\treeHeight(u,v) & \text{if}\,\, u\neq v
\end{cases}\quad \forall u,v\in V\\
\mathrm{where}& \quad  M \equiv \max_{u',v'\in V,\,u'\neq v'}\treeHeight(u',v').
\end{align}
By definition, we have that  $d_{T^*}: V \times V \rightarrow \mathbb{R}^{+}$, $d_{T^*}(u,v)=d_{T^*}(v,u)$, and $d_{T^*}(u,u)= 0$. Thus, to prove that $(V, d_{T^*})$ is an ultrametric, we only need to prove that
\begin{equation}
\interact{}_{T^*}(u,v) \geq \min \{\interact{}_{T^*}(u,t), \interact{}_{T^*}(v,t)\}
\end{equation}
 for every distinct $u,v,t\in V$. In other words, given two leaves nodes $u,v\in V$, the depth $\interact{}_{T^*}(u,v)$ associated to their least common ancestor $i\in T^*$ should be both lower than the depth of any $i$-descendant node in $T^*$ and higher than any $i$-ancestor node in $T^*$:
 % that have the most positive interaction   two clusters merged by the algorithm will always have an interaction that is more negative as compared to all clusters already merged and more positive than all clusters that will be merged at later iterations of the algorithm.
\begin{equation}
\interact{}_{T^*}(u,v) \geq \interact{}_{T^*}(u,t) \Leftrightarrow T^*[u \vee t] \subseteq T^*[u \vee v] 
\end{equation}
 This happens only if, at every iteration, Algorithm \ref{main_alg} merges pairs of clusters with lower and lower interactions, which is true for only certain choices of linkage criteria. %In the next section \ref{sec:alg_update_rules}, we will show that this happens only for certain linkages and only when constraints are not enforced

@@

\TODO{Specify greediness wrt multicut energy, ultrametric properties, and shift-invariance properties} Our main contribution is the generalized algorithm for signed graph partitioning, GASP, that encompasses several known and new agglomerative algorithms on display in Table \ref{tab:linkage-criteria}.
In our framework, individual algorithms are differentiated by the linkage criterion employed. We review them in the following paragraphs.

@@

% \begin{figure}[t]
% \centering
%         \begin{subfigure}[t]{0.46 \textwidth}
%         \centering
%         \includegraphics[width=\textwidth]{figs/example_no_constr.pdf}
%         \caption{No constraints}\label{subfig:no_constraints}
%     \end{subfigure} \hfill \vspace{8pt}
%     \begin{subfigure}[t]{0.46 \textwidth}
%         \centering
%         \includegraphics[width=\textwidth]{figs/example_with_constr.pdf}
%         \caption{With cannot-link constraints}\label{subfig:with_constraints}
%     \end{subfigure}
% \caption{Some iterations of the generalized algorithm (using \emph{Sum} linkage criteria) with and without adding cannot-link constraints. The graph has both attractive (green) and repulsive (red) edges and cannot-link constraints are shown with triple violet bars on the edges. The edge selected at each iteration is highlighted in yellow. We note that when constraints are enforced, the final clustering is given by two clusters instead of only one.}
% \label{fig:algorithm_with_without_CLC}
% \end{figure}

@@

            \thead{\textbf{Unsigned Graphs}} & \thead{HC-Sum-Link} &\thead{HC-Single-Link\\$(\dagger,\ddagger)$} &\thead{HC-Avg-Link\\$(\dagger,\ddagger)$} &\thead{HC-Single-Link\\$(\dagger,\ddagger)$} &\thead{HC-Complete-Link\\$(\dagger,\ddagger)$} \\
            \thead{\textbf{Signed Graphs} \\\textbf{Without Constraints}} & \thead{GAEC \cite{keuper2015efficient}} & \thead{Mutex Watershed \cite{wolf2018mutex}\\$(\dagger)$}& \thead{HC-Avg-Link\\$(\dagger,\ddagger)$} &\thead{HC-Single-Link\\$(\dagger,\ddagger)$} &\thead{HC-Complete-Link\\$(\dagger,\ddagger)$} \\
            \thead{\textbf{Signed Graphs} \\\textbf{With Constraints}} & \thead{Constr-Sum-Link\\(\textbf{NEW})} % \thead{Greedy Fixation \cite{levinkov2017comparative}} 
            & \thead{Mutex Watershed \cite{wolf2018mutex}\\$(\dagger)$}& \thead{Constr-Avg-Link\\(\textbf{NEW})} &  \thead{Constr-Single-Link\\(\textbf{NEW})} &  \thead{Constr-Complete-Link\\(\textbf{NEW})} \\
            % \multicolumn{2}{c|}{\multirow{2}{*}[-0.5em]{\thead{\textbf{\algname{} linkage criteria} $\,\,\interact(S_u ,S_v)$}}}  & \multirow{2}{*}[-0.5em]{\thead{\textbf{Unsigned Graphs}}} & \multicolumn{2}{c}{\thead{\textbf{Signed Graphs}}}  \\        
            % \multicolumn{2}{c|}{} &  &  \multicolumn{1}{c}{\thead{No Constraints}} & \thead{With Constraints} \\ \midrule

@@

        \end{tabular}
    \end{subtable} 
    \caption{Existing and new clustering algorithms that can be reformulated as special cases of the proposed generalized algorithm for signed graph partitioning, \algname{}, given a linkage criterion, a type of graph (signed or unsigned) and the optional use of cannot-link constraints. The set $E_{uv}$ is defined as the set of all edges connecting cluster $S_u$ to cluster $S_v$, i.e. $E_{uv}=(S_u \times S_{v \neq u}) \cap E$. \TODO{$\dagger$ if the linkage defines ultrametric and $\ddagger$ if it defines a weight-shift invariant algorithm} } 
    \label{tab:linkage-criteria}
\end{table*}
